Fish_Stuff

### 2024-08-20 R/Fish_ord.Rmd
ordinated with overall herb biomass and functional biomass but nothing explains much :fish:

```{r data, message=FALSE, echo=FALSE}
#algae <- read.csv(here("data", "algaeonly_cleaned.csv"))
fish_fun <- read.csv(here("data", "biomass_functional_summary.csv"))
fish <- read.csv(here("data", "biomass_herb_summary.csv"))
alpha <- read.csv(here("data", "algae_alpha_diversity.csv")) %>% 
  filter(Habitat != "Outer 17")
```

@@

### Ordination of Algal Communities with Fish Community as Drivers:

@@

```{r}
fish_fun1 <- fish_fun
fish_fun1[is.na(fish_fun1)] <- 0
fish_fun1 <- fish_fun1 %>% 
  mutate(Site = paste0("LTER ",Site))
merged <- left_join(wide_site_level_merge, fish_fun1, by = join_by(Year, Site, Habitat))
merged_clean <- merged[complete.cases(merged[c("Browser", "Brusher", "Cropper", "Concealed.Cropper", "Excavator", "Scraper")]), ]

>>>>

```{r data, message=FALSE, echo=FALSE}
#algae <- read.csv(here("data", "algaeonly_cleaned.csv"))
fish_fun <- read.csv(here("data", "biomass_functional_summary.csv"))
herb <- read.csv(here("data", "biomass_herb_summary.csv"))
alpha <- read.csv(here("data", "algae_alpha_diversity.csv")) %>% 
  filter(Habitat != "Outer 17")
```

@@

# Ordination of Algal Communities with Fish Community as Drivers:

@@

## Just Herbivore Biomass:
```{r}
herb_merge <- herb %>% 
  mutate(Site = paste0("LTER ",Site))
herb_algal_merge <- left_join(wide_site_level_merge, herb_merge[,c("Year", "Site", "Habitat", "Herbivore.Biomass")], 
                              by = join_by(Year, Site, Habitat)) %>% 
  filter(Year != "2023")

species.data.transform = wisconsin(sqrt(herb_algal_merge[c(5:74)]))
non_zero_rows <- rowSums(species.data.transform == 0) != ncol(species.data.transform)
species.data.use <- species.data.transform[non_zero_rows, ]
merged_clean_filtered <- herb_algal_merge[non_zero_rows, ]
```

```{r}
herb_algal_constrained_ord = capscale(species.data.use ~ Herbivore.Biomass, dist = "bray", data = merged_clean_filtered)
```

```{r}
herb_algal_constrained_ord
```

# Plot 1: Species driving the community composition
plot(herb_algal_constrained_ord, display = "species", main = "Species Driving Community Composition", type = "n")
text(herb_algal_constrained_ord, display = "species", col = "blue", cex = 1)

# Plot 2: Communities (sites) plotted
plot(herb_algal_constrained_ord, display = "sites", main = "Communities", type = "n")
points(herb_algal_constrained_ord, display = "sites", col = color_palette[colors], pch = 21)  # Sites colored by the same scheme

# Plot 3: Constrained ordination axes (environmental variables)
plot(herb_algal_constrained_ord, display = "sites", type = 'n', main = "Relationship with Herbivore Biomass")
points(herb_algal_constrained_ord, display = "sites", col = 'grey')
text(herb_algal_constrained_ord, display = 'bp', col = 'blue')
```

## The whole fish community:
```{r}
fish_fun1 <- fish_fun
fish_fun1[is.na(fish_fun1)] <- 0
fish_fun1 <- fish_fun1 %>% 
  mutate(Site = paste0("LTER ",Site))
merged <- left_join(wide_site_level_merge, fish_fun1, by = join_by(Year, Site, Habitat))
merged_clean <- merged[complete.cases(merged[c("Browser", "Brusher", "Cropper", "Concealed.Cropper", "Excavator", "Scraper")]), ]
